---
title: "Lab 01"
author: "Cristian E. Nuno"
date: "September 3, 2016"
output: pdf_document
---
```{r, echo=F, warning=FALSE }
# Load Necessary Packages and Data
library( Lahman )
data( Master )
```


## Question 1

What is the average weight of an MLB player?

Answer: Approximately 186 pounds.
```{r}
mean(Master$weight, na.rm=TRUE)
```
## Question 2

How many MLB players have come from the Netherlands?

Answer: 12
```{r}
a <- table( Master$birthCountry )
a [names(a)=="Netherlands"]
```
## Question 3

How tall was the tallest player in MLB? The shortest?

Answer: The tallest player was 83 inches (nearly 7 feet) tall; meanwhile, the shortest player was 43 inches (shorter than 4 feet) tall.
```{r}
max(Master$height, na.rm=TRUE)
min(Master$height, na.rm=TRUE)
```
## Question 4

What percentage of players in MLB switch-hit?

Answer: 6.59 perecent of players in MLB can switch-hit.
```{r}
tbl <- table(Master$bats)
res <- cbind(tbl,round (prop.table(tbl) * 100, 2))
colnames(res) <- c('Count', 'Percentage')
res
```
## Question 5

How many players are right-handed when they throw, but left-handed when they bat?

Answer: 2054 players in MLB throw with their right-hand but are left-handed when they bat.
```{r}
table(Master$throws, Master$bats)
```

