---
output: html_document
runtime: shiny
---
---
title: "Visualizing City of Cincinnati Monthly Citizen Complaints Against Police Officers "
author: "Cristian E. Nuno"
date: "2/1/2017"
runtime: shiny
output: html_document
---
```{r, echo=FALSE, message=FALSE, fig.height=6, fig.width=9}
# Load Data
library( dplyr )
ccc_url <- "https://data.cincinnati-oh.gov/api/views/r3vg-n6p3/rows.csv?accessType=DOWNLOAD"
ccc <- read.csv( ccc_url, stringsAsFactors = FALSE)

# Convert Date.Received class from character to date class
ccc$Date.Received <- as.Date( ccc$Date.Received, format = "%m/%d/%Y")

# Aggregate by year and by month
# Plot aggregate total citizen complaints over time
year.month.all <- format( ccc$Date.Received, 
                      format = "%Y-%m" )

# Create quartiles from frequency of complaints
# 3rd quartile value is 14
quantiles.freq <- as.data.frame( quantile( 
                                  table( year.month.all ) 
                                  ) 
                                 )
# Select and color those frequencies which are above the 75th quartile
# (FALSE + 1 -> 1 -> "blue";    TRUE + 1 -> 2 -> "red")
above.75th <- c( "#002673", "#C70F2E")[ ( table( year.month.all ) > quantiles.freq$Freq[ 4 ] ) + 1 ]

# Variables that may come in handy later
as.month <- format( ccc$Date.Received, format = "%b")
month.order <- c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec")
as.month <- factor( as.month, ordered = TRUE, levels = month.order )
as.year <- format( ccc$Date.Received, format = "%Y" )
ccc$Year <- as.year
ccc$Month <- as.month

# Aggregated complaints by neighborhood
n.count <- as.data.frame( table( ccc$Neighborhood ) )
colnames( n.count ) <- c("Neighborhood", "Freq")

# Determine values that are above the 3rd Quartile;
# 3rd Quartile = specific value where 75% of values fall below that specified value

# 18.25 is our value
#quantile( n.count$Freq ) 
# Identify neighborhoods who have complaints over 18.25
high.n.count <- which( n.count$Freq > 18.25 ) 
# List these neighborhoods
# as.character( n.count$Neighborhood[ high.n.count ] )
month.labels <- c("JAN", "FEB", "MAR", "APR", "MAY", "JUN", "JUL", "AUG", "SEP", "OCT", "NOV", "DEC")

```

## The monthly complaints filed by citizens against the City of Cincinnati police officers database was recently made publicly available:
https://data.cincinnati-oh.gov/Safer-Streets/Citizen-Complaints/r3vg-n6p3/data.

### At a glance, click below for the total monthly complaints for all of Cincinnati from 2010 to 2015:
https://github.com/cenuno/MPA_Portfolio/blob/lab_01/cenuno_Cin_Citz_Complaints_3.pdf



## Below is an interactive tool which visualizes how many monthly complaints were filed in their neighborhood from 2010 to 2015.

```{r, echo = FALSE}
# Now we are going to create an interactive visual
# Where the multiplot created will depend on 
# The user to select the neighborhood as an
# Input value.

# Remember these are the neighborhoods that have...
# more than 18.25 complaints from 2010-2015:
# "Avondale"                   "Central Business Section"  
# "Clifton-University Heights" "Corryville"                
# "East Price Hill"            "Evanston"                  
# "Over-The-Rhine"             "Walnut Hills"              
# "West End"                   "West Price Hill"           
# "Westwood"                   "Winton Hills"

library( shiny )

selectInput( inputId = "neighborhood", 
            label = "Select a Cincinnati neighborhood to examine:", 
            choices = as.character( unique( n.count$Neighborhood ) ),
            selected = "Avondale"
          )


renderPlot({
# Plot user input neighborhood
# Filter data by neighborhood the user selects, resulting in 7 filtered datasets
user.year.all <-  filter( ccc, Date.Received >= "2010-01-01" &
                               Date.Received <= "2015-12-31" &
                               Neighborhood == input$neighborhood )

user.year.2010 <- filter( ccc, Date.Received >= "2010-01-01" &
                                Date.Received <= "2010-12-31" & 
                                Neighborhood == input$neighborhood )

user.year.2011 <- filter( ccc, Date.Received >= "2011-01-01" &
                               Date.Received <= "2011-12-31" &
                               Neighborhood == input$neighborhood )

user.year.2012 <- filter( ccc, Date.Received >= "2012-01-01" &
                               Date.Received <= "2012-12-31" &
                               Neighborhood == input$neighborhood )

user.year.2013 <- filter( ccc, Date.Received >= "2013-01-01" &
                               Date.Received <= "2013-12-31" &
                               Neighborhood == input$neighborhood )

user.year.2014 <- filter( ccc, Date.Received >= "2014-01-01" &
                               Date.Received <= "2014-12-31" &
                               Neighborhood == input$neighborhood )

user.year.2015 <- filter( ccc, Date.Received >= "2015-01-01" &
                               Date.Received <= "2015-12-31" &
                               Neighborhood == input$neighborhood )

# Arrange filtered datasets by the year and month
user.year.month <- format( user.year.all$Date.Received,
                               format = "%Y-%m")

user.year.month.2010 <- format( user.year.2010$Date.Received,
                                    format = "%Y-%m" )

user.year.month.2011 <- format( user.year.2011$Date.Received,
                                    format = "%Y-%m" )

user.year.month.2012 <- format( user.year.2012$Date.Received,
                                    format = "%Y-%m" )

user.year.month.2013 <- format( user.year.2013$Date.Received, 
                                    format = "%Y-%m" )

user.year.month.2014 <- format( user.year.2014$Date.Received, 
                                    format = "%Y-%m" )

user.year.month.2015 <- format( user.year.2015$Date.Received, 
                                    format = "%Y-%m" )
# Plot User Input Chart
# Need to scale each graph on a similar plot (total for user input neighborhood)
y.max.user <- max( table( user.year.month ) )

# Need to make sure each month is accounted for, even when the instance is 0.
levels.2010 = c( "2010-01", "2010-02", "2010-03", "2010-04", 
                 "2010-05", "2010-06", "2010-07", "2010-08", 
                 "2010-09", "2010-10", "2010-11", "2010-12")

levels.2011 = c( "2011-01", "2011-02", "2011-03", "2011-04", 
                 "2011-05", "2011-06", "2011-07", "2011-08", 
                 "2011-09", "2011-10", "2011-11", "2011-12")

levels.2012 = c( "2012-01", "2012-02", "2012-03", "2012-04", 
                 "2012-05", "2012-06", "2012-07", "2012-08", 
                 "2012-09", "2012-10", "2012-11", "2012-12")

levels.2013 = c( "2013-01", "2013-02", "2013-03", "2013-04", 
                 "2013-05", "2013-06", "2013-07", "2013-08", 
                 "2013-09", "2013-10", "2013-11", "2013-12")

levels.2014 = c( "2014-01", "2014-02", "2014-03", "2014-04", 
                 "2014-05", "2014-06", "2014-07", "2014-08", 
                 "2014-09", "2014-10", "2014-11", "2014-12")

levels.2015 = c( "2015-01", "2015-02", "2015-03", "2015-04", 
                 "2015-05", "2015-06", "2015-07", "2015-08", 
                 "2015-09", "2015-10", "2015-11", "2015-12")

# arrange plots on one space, 2 x 3
par( mfrow = c(2, 3), oma = c(3.5, 0, 4, 0) ) 

# Plot 2010 user input data
plot( table( factor( user.year.month.2010, levels = levels.2010 ) ), 
      ylim = c( 0, y.max.user), 
      lty = 1, lwd = 5, cex.axis = 0.8, 
      col = "#002673", 
      ylab = "", yaxt = "n",
      xlab = "", xaxt = "n" )
# Add x-axis
axis( side = 1, at=seq(1, 12, by=1), labels = month.labels, 
      cex.axis = 1, lwd = 1, mgp = c(0,0,1), 
      col.ticks = "white", col.axis = gray(0.5), col = "white")
# Add y-axis
axis(side=2, at=seq(0, y.max.user, by=5 ), 
     las=1, cex.axis=1,
     col.axis=gray(.5), col="white" )
# Add title
mtext( side = 3, line = 1, cex = 1.2, adj = 0.5,
       text = "2010")

# Plot 2011 user data
plot( table( factor( user.year.month.2011, levels = levels.2011 ) ), 
      ylim = c( 0, y.max.user), 
      lty = 1, lwd = 5, cex.axis = 0.8, 
      col = "#002673", 
      ylab = "", yaxt = "n",
      xlab = "", xaxt = "n" )
# Add x-axis
axis( side = 1, at=seq(1, 12, by=1), labels = month.labels, 
      cex.axis = 1, lwd = 1, mgp = c(0,0,1), 
      col.ticks = "white", col.axis = gray(0.5), col = "white")
# Add y-axis
axis(side=2, at=seq(0, y.max.user, by=5 ), 
     las=1, cex.axis=1,
     col.axis=gray(.5), col="white" )
# Add title
mtext( side = 3, line = 1, cex = 1.2, adj = 0.5,
       text = "2011")

# Plot 2012 user data
plot( table( factor( user.year.month.2012, levels = levels.2012 ) ), 
      ylim = c( 0, y.max.user), 
      lty = 1, lwd = 5, cex.axis = 0.8, 
      col = "#002673", 
      ylab = "", yaxt = "n",
      xlab = "", xaxt = "n" )
# Add x-axis
axis( side = 1, at=seq(1, 12, by=1), labels = month.labels, 
      cex.axis = 1, lwd = 1, mgp = c(0,0,1), 
      col.ticks = "white", col.axis = gray(0.5), col = "white")
# Add y-axis
axis(side=2, at=seq(0, y.max.user, by=5 ), 
     las=1, cex.axis=1,
     col.axis=gray(.5), col="white" )
# Add title
mtext( side = 3, line = 1, cex = 1.2, adj = 0.5,
       text = "2012")

# Plot 2013 user data
plot( table( factor( user.year.month.2013, levels = levels.2013 ) ), 
      ylim = c( 0, y.max.user), 
      lty = 1, lwd = 5, cex.axis = 0.8, 
      col = "#002673", 
      ylab = "", yaxt = "n",
      xlab = "", xaxt = "n" )
# Add x-axis
axis( side = 1, at=seq(1, 12, by=1), labels = month.labels, 
      cex.axis = 1, lwd = 1, mgp = c(0,0,1), 
      col.ticks = "white", col.axis = gray(0.5), col = "white")
# Add y-axis
axis(side=2, at=seq(0, y.max.user, by=5 ), 
     las=1, cex.axis=1,
     col.axis=gray(.5), col="white" )

# Add title
mtext( side = 3, line = 1, cex = 1.2, adj = 0.5,
       text = "2013")

# Plot 2014 user data
plot( table( factor( user.year.month.2014, levels = levels.2014 ) ), 
      ylim = c( 0, y.max.user), 
      lty = 1, lwd = 5, cex.axis = 0.8, 
      col = "#002673", 
      ylab = "", yaxt = "n",
      xlab = "", xaxt = "n" )
# Add x-axis
axis( side = 1, at=seq(1, 12, by=1), labels = month.labels, 
      cex.axis = 1, lwd = 1, mgp = c(0,0,1), 
      col.ticks = "white", col.axis = gray(0.5), col = "white")
# Add y-axis
axis(side=2, at=seq(0, y.max.user, by=5 ), 
     las=1, cex.axis=1,
     col.axis=gray(.5), col="white" )
# Add title
mtext( side = 3, line = 1, cex = 1.2, adj = 0.5,
       text = "2014")

# Plot 2015 all data
plot( table( factor( user.year.month.2015, levels = levels.2015 ) ), 
      ylim = c( 0, y.max.user), 
      lty = 1, lwd = 5, cex.axis = 0.8, 
      col = "#002673", 
      ylab = "", yaxt = "n",
      xlab = "", xaxt = "n" )
# Add x-axis
axis( side = 1, at=seq(1, 12, by=1), labels = month.labels, 
      cex.axis = 1, lwd = 1, mgp = c(0,0,1), 
      col.ticks = "white", col.axis = gray(0.5), col = "white")
# Add y-axis
axis(side=2, at=seq(0, y.max.user, by=5 ), 
     las=1, cex.axis=1,
     col.axis=gray(.5), col="white" )
# Add title
mtext( side = 3, line = 1, cex = 1.2, adj = 0.5,
       text = "2015")

# Title the entire multiplot
n.name <- unique( user.year.all$Neighborhood )
n.name <- paste( n.name, "Monthly Complaints Filed Against\nCincinnati Police Officers, 2010-2015", 
                 sep = " " )
mtext( text = n.name, outer = TRUE, cex = 1.2 )

# Add date created, who made it, and data source
mtext( outer = TRUE, cex = 0.6, adj = 0.98, 
       line = 0.25, side = 1,
       text = "Date Created: 02/01/2017")

mtext( outer = TRUE, cex = 0.6, adj = 0.98,
       line = 1, side = 1,
       text = "Created By: Cristian E. Nuno")

mtext( outer = TRUE, cex = 0.6, adj = 0.98, 
       line = 1.75, side = 1,
       text = "Date Source: https://data.cincinnati-oh.gov/Safer-Streets/Citizen-Complaints/r3vg-n6p3/data")

})

```
### Here are summary statistics regarding each neighborhood's total number of monthly complaints, 2010-2015:

```{r, echo = FALSE, message = FALSE}
library( pander )
pander( table( ccc$Neighborhood ))
```

###These are the 12 neighborhoods that filed more than 18 complaints from 2010-2015:
##### *18 is the total number of complaints which splits off the highest 25% of data from the lowest 75%.*

* Avondale  
* Central Business Section   
* Clifton-University Heights  
* Corryville          
* East Price Hill         
* Evanston               
* Over-The-Rhine           
* Walnut Hills              
* West End                
* West Price Hill           
* Westwood                  
* Winton Hills