---
title: "Lab 03"
author: "Cristian E. Nuno"
date: "September 19, 2016"
output: html_document
---
```{r results='hide', message=FALSE}
# Load Necessary Packages and Data
library( Lahman )
data( Master )
data( Teams )
data( Salaries )
data( HallOfFame )
data( Schools )
data( CollegePlaying )

# For table formatting purposes
library( pander )
library( plyr )
library( dplyr )
```


## Question 1

What is the average salary of a player on the Oakland A’s in 2002? A player on the Boston Red Sox that same year? What is the total budget of each franchise in 2002? 

Answer: The median salary of a player on the Oakland A's in 2002 was $812,500; meanwhile, the total budget of the team was just over $40 million dollars. For the same year, the median salary of a player on the Boston Red Sox was $2,287,500; however, their total budget was over $100 million. 

```{r}
# Identify shared variable names between the two datasets
intersect( names( Teams ), names( Salaries ) )


# Merge the "Teams" and "Salaries" datasets
trial <- merge( Teams, Salaries, by.x=c( "teamID", 
                                         "yearID", 
                                         "lgID" ), 
                                 by.y=c( "teamID", 
                                         "yearID", 
                                         "lgID" )
                )


# Median salary in 2002 of Oakland A's players
oak <-  trial[ trial$teamID == "OAK" & trial$yearID == 2002, ]
os <- median( oak$salary )
paste( "Median 2002 Oakland A's Salary = $", format( os, big.mark="," ) , sep="" )

# Total budget Oakland A's in 2002

tos <- sum( oak$salary )
paste( "Total 2002 budget for Oakland A's = $", format( tos, big.mark="," ) , sep="" )

# Median salary in 2002 of Boston Red Sox players
bss <-  trial[ trial$teamID == "BOS" & trial$yearID == 2002, ]
bs <- median( bss$salary )
paste( "Median 2002 Boston Red Sox Salary = $", format( bs, big.mark="," ) , sep="" )

# Total budget Boston Red Sox in 2002
tbs <- sum( bss$salary )
paste( "Total 2002 budget for Boston Red Sox = $", format( tbs, big.mark="," ) , sep="" )
```

## Question 2

Which city (birth city) has produced the most baseball hall-of-famers? Use the Master and HallOfFame datasets.

Answer: Chicago has produced the most baseball hall-of-famers at 9 as of today.
```{r}
intersect( names(HallOfFame), names(Master) )
HOF.Cities <- merge( HallOfFame, Master, by.x="playerID", by.y="playerID" )
hofwinners <- HOF.Cities[ HOF.Cities$inducted == "Y", ]
which.max( table( hofwinners$birthCity ) )
table( hofwinners$birthCity ) [45]
```

## Question 3

How many players went to college in Los Angeles, San Diego, or San Francisco? Use the CollegePlaying and the Schools datasets.

Answer: 347 players went to college in either Los Angeles, San Diego, or San Francisco.
```{r}
dat.schools <- merge( CollegePlaying, Schools )
head( dat.schools )
head( sort( table( dat.schools$city ), decreasing=T ) )

dat.schools <- dat.schools[ , c("playerID","name_full","city") ]
nrow( dat.schools )

dat.schools <- unique( dat.schools )
nrow( dat.schools )

head( sort( table( dat.schools$city ), decreasing=T ) )
sum( dat.schools$city == "Los Angeles" |
     dat.schools$city == "San Diego" |
     dat.schools$city == "San Francisco" )
```

