---
title: "Lab_02"
author: "Cristian E. Nuno"
date: "September 13, 2016"
output: html_document
---
```{r Load Required Packages and Data}
library( pander )
library( dplyr )
dat <- read.csv( "https://raw.githubusercontent.com/lecy/maps-in-R/master/Data/syr_parcels.csv" )
```

## Question 1

What is the total assessed value of all the taxable land in Syracuse?

Answer: $ 7,538,424,950
```{r}
total.AV <- sum( dat$AssessedVa, na.rm=TRUE )
paste( "$", format( total.AV, big.mark="," ), sep="" )
```


## Question 2

What proportion of land parcels is not taxable? What is the value of these non-taxable parcels?

Answer: The proportion of land parcels that is not taxable is approximately 12 percent of all parcels, but approxmiately 40 percent of all assessed parcels. The value of assessed non-taxable parcels is $3,006,860,002.
```{r}
# Proportion of land parecels that is not taxable
nontaxparcels <- dat$LandUse == "Cemetery" | dat$LandUse == "Community Services" | 
                 dat$LandUse == "Parking" | dat$LandUse == "Parks" | dat$LandUse == "Recreation" |
                 dat$LandUse == "Religious" | dat$LandUse == "Schools" |
                 dat$LandUse == "Utilities" | dat$LandUse == "Vacant Land"

# Sum of these nontaxable parcels
sum( nontaxparcels )

# Identifying taxable parcels
taxparcels <- ! nontaxparcels

# Calculating proportion of nontaxable parcels
propnontax <- sum( nontaxparcels ) / sum( nontaxparcels, taxparcels)
paste(round(100*propnontax, 2), "%", sep="")

# Value of assessed nontaxable parcels
Value <- sum( dat$AssessedVa [ nontaxparcels ] )
paste( "$", format( Value, big.mark="," ), sep="" )

# Calculating proportion of assessed nontaxable parcels
Proportion <- Value / sum( dat$AssessedVa )
paste( round(100*Proportion, 2), "%", sep="" )
```


## Question 3

Which property has the highest delinquent tax bill?

Answer: The James M. Hanley Federal Building, owned by the U.S. federal government, is the property with the highest delinquent. The total amount of the delinquient tax bill was $1,543,354.
```{r}
max.del.tbill <- which.max( dat$AmtDelinqu )
dat[ max.del.tbill, ] %>% pander
```


## Question 4

Which neighborhoods have the highest proportion of vacant buildings?

Answer:

1.  Brighton at 11%.

2.  Near Westside & Prospect Hill at 10%.

3.  Elmwood & Southside at 9%.

4.  Northside & Southwest at 8%.

```{r}
table( dat$Nhood, dat$VacantBuil ) %>% prop.table( margin=1 ) %>% round( 2 ) %>% pander
```


## Question 5

How many single family homes worth more than $100,000 each are located in each of the neighborhoods above?

Answer:

*  Brighton has 0.

*  Near Westside & Prospect HIll each have 1.

*  Elmwood has 6. Southside has 0.

*  Northside has 5. Southwest has 0.
```{r}
# Produce a table of all neighborhoods
these.single.fam.100k <- dat$LandUse == "Single Family" & dat$AssessedVa > 100000
table( dat$Nhood, these.single.fam.100k ) %>% pander
```

