---
title: "Arbolibro: A Community Book-Share Project"
author: ""
output:
  html_document:
    css: dc_style.css
    toc: true
    toc_float: true
---
# Our Mission
<span style="color:#3399ff">
<font size="5">
Arbolibro's mission is for children and adults to take books home, read, and enjoy them for free. 
</font>
</span>

```{r, echo=FALSE, fig.height=4, fig.width=4, fig.align='center'}
# image packages
#<<<<<<< HEAD
#=======
# putting this inside Arbolibro project tryingß
# no idea what this is>>>>>>>
# 4e6854157bc31b2e61714874422ac8ee3b5c37ff
library(png)
library(grid)
img <- readPNG('/Users/cristiannuno/Desktop/Syracuse/book1.png')
grid.raster(img)
```

</center>

# Our History

<span style="color:#3399ff">
<font size="5">
Arbolibro was born in August of 2016 by me, thanks to my kids and their love of reading. 
</font>
</span>

```{r, echo=FALSE, fig.height=4, fig.width=4, fig.align='center'}
# image packages
library(png)
library(grid)
img <- readPNG('/Users/cristiannuno/Desktop/Syracuse/book2.png')
grid.raster(img)
```

<span style="color:#3399ff">
<font size="5">
Working and going to school full-time only allowed once a week trips to the local library. One day, I thought of the idea of starting a book-share project in my Little Village neighborhood in Chicago, IL. Now our kids can grab a book right down the street instead of waiting for me to take them to the library.
</font>
</span>


