---
output:
  html_document:
    css: dc_style.css
    toc: true
    toc_float: true
---
# ArboLibro: A Community Book-Share Project

```{r, echo=FALSE, fig.height=8, fig.width=12, fig.align='center'}
# image packages
library(png)
library(grid)
img <- readPNG('/Users/cristiannuno/Desktop/Syracuse/Arbolibro_Logo.png')
grid.raster(img)
```


## Our Mission
<span style="color:#DC143C">
<font size="5">
ArboLibro's mission is for children and adults to take books home, read, and enjoy them for free. 
</font>
</span>

```{r, echo=FALSE, fig.height=6, fig.width=6, fig.align='center'}
# image packages
library(png)
library(grid)
img <- readPNG('/Users/cristiannuno/Desktop/Syracuse/book1.png')
grid.raster(img)
```

## Our History

<span style="color:#DC143C">
<font size="5">
ArboLibro was born in August of 2016 thanks to my kids and their love of reading. 
</font>
</span>

```{r, echo=FALSE, fig.height=6, fig.width=6, fig.align='center'}
# image packages
library(png)
library(grid)
img <- readPNG('/Users/cristiannuno/Desktop/Syracuse/book2.png')
grid.raster(img)
```

<span style="color:#DC143C">
<font size="5">
Working and going to school full-time only allowed once a week trips to the local library. One day, I thought of the idea of starting a book-share project in my Little Village neighborhood in Chicago, IL. Now our kids can grab a book right down the street instead of waiting for me to take them to the library.
</font>
</span>

## Our Motivation

<span style="color:#DC143C">
<font size="5">
Families from low-income communities struggle with literacy for several reasons. Limited education, a need to work multiple jobs, and little knowledge or access to resources. ArboLibro **empowers and educates** the community by offering free literature for all ages.
</font>
</span>


<span style="color:#DC143C">
<font size="5">
In the United States, book-sharing has been promoted as a valuable activity that parents and children can do together to enhance literacy skills. Reading with your child has many positive benefits and is the key to your child’s academic future. 
</font>
</span>

```{r, echo=FALSE, fig.height=6, fig.width=6, fig.align='center'}
# image packages
library(png)
library(grid)
img <- readPNG('/Users/cristiannuno/Desktop/Syracuse/book3.png')
grid.raster(img)
```

<span style="color:#DC143C">
<font size="5">
It benefits children’s neurological, educational, psychological, social and linguistic future. Arbolibro offers a wide range of books from all ages in both English and Spanish. This project will bring an alternative to the community bringing them together.
</font>
</span>


# Arbolibro Locations

<span style="color:#DC143C">
<font size="5">
We currently have 4 locations in the neighborhoods ofLittle Village and Pilsen. ArboLibro strives to build in neighborhoods which are historically low-income. We are looking forward to expanding to neighborhoods throughout Chicago in the near future.
</font>
</span>



```{r Necessary Packages, echo=FALSE, message=FALSE}
library( leaflet )
library( magrittr )
library( readxl )
library( rgdal )
library( stringi )
library( lucr )
```



```{r Import Data, echo=FALSE, results='hide'}
# Census Data - Selected socioeconomic indicators in Chicago, 2008 – 2012
dat <- read.csv( "https://data.cityofchicago.org/api/views/kn9c-c2s2/rows.csv?accessType=DOWNLOAD"
                 , header = TRUE
                 , stringsAsFactors = FALSE)
# The last row in this dataset is a row representing the entire city
city.wide.info <- tail( dat, n = 1)

# Remove the last row from the dataset
dat <- dat[ -nrow( dat ), ]

# City of Chicago Community Area Boundaries in GeoJSON format
chicago.com.area <- readOGR("https://data.cityofchicago.org/api/geospatial/cauq-8yn6?method=export&format=GeoJSON"
                            , "OGRGeoJSON"
                            , stringsAsFactors = FALSE
)
# Transform spatial polygon data frame's "community area number"" variable
# from character to integer
chicago.com.area@data$area_numbe <- as.integer( chicago.com.area@data$area_numbe )
chicago.com.area@data$area_num_1 <- as.integer( chicago.com.area@data$area_num_1 )

chicago.com.area@data <- data.frame( chicago.com.area@data
                                     , dat[
                                       match( chicago.com.area@data[ , "area_numbe" ]
                                              , dat[ , "Community.Area.Number"  ]
                                       )
                                       , ]
)

arbolibro <- read_excel( path = "/Users/cristiannuno/Desktop/Syracuse/arbolibro.xlsx"
                   , sheet = 1
                   )
# Transform location urls from static to active
arbolibro$ActiveURL <- paste0("<a href='"
                                   , arbolibro$LocationWebsite
                                   , "' target='_blank'>"
                                   , arbolibro$LocationWebsite
                                   ,"</a>"
                                   )
```



```{r Design Map, echo=FALSE, fig.height=4, fig.width=8, fig.align='center'}
# Design a color palette for the map
pal <- colorNumeric(
  palette = "YlGnBu",
  domain = NULL #chicago.com.area$PER.CAPITA.INCOME # vector 
)

# Build Map
mapArbolibro <- leaflet( chicago.com.area ) %>% # add locations dataset
  setView( lng = -87.705823, lat = 41.845816, zoom = 12) %>% # Set map to Little Village neighborhood
  addTiles() %>%  # Add map background
  addMarkers( data = arbolibro, lng = ~Long, lat = ~Lat
              # Add useful information onto each marker's popup
              , popup = paste0( "<b> Location Name: </b>"
                                , arbolibro$LocationName
                                , "<br>"
                                , "<b>Location Address: </b>"
                                , arbolibro$Location
                                , "<br>"
                                , "<b> Location Type: </b>"
                                , arbolibro$Type
                                , "<br>"
                                , "<b> Location Website: </b>"
                                , arbolibro$ActiveURL
                                )
              ) %>%
  addPolygons( smoothFactor = 0.2
              , fillOpacity = 0.7
              # Color the polygons based on the value of their
              # per capita income
              , color = ~pal(PER.CAPITA.INCOME)
              # Add useful information onto the popup
              , popup = paste0( "<b>Community Area Name: </b>", stri_trans_totitle( chicago.com.area@data$community ), "<br>"
                                , "<b>Community Area Number: </b>", chicago.com.area@data$area_numbe, "<br>"
                                , "<b>Per Capita Income: </b>", to_currency( chicago.com.area$PER.CAPITA.INCOME
                                                                             , currency_symbol = "$"
                                                                             )
                                )
              # Whenever someone hovers their mouse above a polygon,
              # make sure they know which community they are in
              , label = chicago.com.area@data$community
              , labelOptions = labelOptions( textsize = "25px", textOnly = TRUE )
              , highlightOptions = highlightOptions( color = "orange"
                                                      , weight = 6
                                                      , bringToFront = TRUE
                                                      )
                
  ) %>%
  # Make sure people know what the color values indicate
  addLegend("bottomleft"
            , pal = pal # use the same color palette we made earlier
            , values = ~PER.CAPITA.INCOME # assign values to the legend
            , title = "Per Capita Income (in 2012 dollars)"
            , labFormat = labelFormat(prefix = "$")
            , opacity = 1
  ) %>%
  addProviderTiles( "OpenStreetMap_DE" )

# Display Map
mapArbolibro

```

# In the News

<span style="color:#DC143C">
<font size="5">
Arbolibro has [received community attention](https://soundcloud.com/vocalo/domingosenvocalo-rayuelafm07316) from Lawndale News and Vocalo 91.1 FM, a segment called Rayuela FM which focuses on literature.
</font>
</span>


<span style="color:#DC143C">
<font size="3">
The interview starts at the 21:15 mark.
</font>
</span>


# Contact Us

<span style="color:#DC143C">
<font size="5">
Please support the Arbolibro project by [recommending where where we should build our next Arbolibro](https://goo.gl/forms/yzEytmYNrLCtumaF2).
</font>
</span>


<span style="color:#DC143C">
<font size="5">
For immediate feedback, please email the founder of Arborlibro, Gonzalo Rizo, at gonzalorizo@hotmail.com. *This page was last updated on April 12, 2017*.
</font>
</span>

## Website Design

<span style="color:#DC143C">
<font size="5">
This website was designed by Cristian E. Nuno. To view more of his work, see his [GitHub page](https://github.com/cenuno) and his [LinkedIn page](https://www.linkedin.com/in/cristiannuno). You may email him at nuno.e.cristian@gmail.com.
</font>
</span>

