---
title: "Arbolibro: A Community Book-Share Project"
author: ""
output:
  html_document:
    css: dc_style.css
    toc: true
    toc_float: true
---
# Our Mission
<span style="color:#3399ff">
<font size="5">
Arbolibro's mission is for children and adults to take books home, read, and enjoy them for free. 
</font>
</span>

```{r, echo=FALSE, fig.height=4, fig.width=4, fig.align='center'}
# image packages
library(png)
library(grid)
img <- readPNG('/Users/cristiannuno/Desktop/Syracuse/book1.png')
grid.raster(img)
```

</center>

# Our History

<span style="color:#3399ff">
<font size="5">
Arbolibro was born in August of 2016 by me, thanks to my kids and their love of reading. 
</font>
</span>

```{r, echo=FALSE, fig.height=4, fig.width=4, fig.align='center'}
# image packages
library(png)
library(grid)
img <- readPNG('/Users/cristiannuno/Desktop/Syracuse/book2.png')
grid.raster(img)
```

<span style="color:#3399ff">
<font size="5">
Working and going to school full-time only allowed once a week trips to the local library. One day, I thought of the idea of starting a book-share project in my Little Village neighborhood in Chicago, IL. Now our kids can grab a book right down the street instead of waiting for me to take them to the library.
</font>
</span>

# Our Motivation

<span style="color:#3399ff">
<font size="5">
Families from low-income communities struggle with literacy for several reasons. Limited education, a need to work multiple jobs, and little knowledge or access to resources. Arbolibro empowers and educates the community offering free literature for all ages.
</font>
</span>


<span style="color:#3399ff">
<font size="5">
In the United States, book-sharing has been promoted as a valuable activity that parents and children can do together to enhance literacy skills. Reading with your child has many positive benefits and is the key to your child’s academic future. 
</font>
</span>


