---
title: "Arbolibro: A Community Book-Share Project"
author: ""
output:
  html_document:
    css: dc_style.css
    toc: true
    toc_float: true
---
# Our Mission
<span style="color:#3399ff">
<font size="5">
Arbolibro's mission is for children and adults to take books home, read, and enjoy them for free. 
</font>
</span>

```{r, echo=FALSE, fig.height=4, fig.width=4, fig.align='center'}
# image packages
library(png)
library(grid)
img <- readPNG('/Users/cristiannuno/Desktop/Syracuse/book1.png')
grid.raster(img)
```

</center>

# Our History

<span style="color:#3399ff">
<font size="5">
Arbolibro was born in August of 2016 by me, thanks to my kids and their love of reading. 
</font>
</span>

```{r, echo=FALSE, fig.height=4, fig.width=4, fig.align='center'}
# image packages
library(png)
library(grid)
img <- readPNG('/Users/cristiannuno/Desktop/Syracuse/book2.png')
grid.raster(img)
```

<span style="color:#3399ff">
<font size="5">
Working and going to school full-time only allowed once a week trips to the local library. One day, I thought of the idea of starting a book-share project in my Little Village neighborhood in Chicago, IL. Now our kids can grab a book right down the street instead of waiting for me to take them to the library.
</font>
</span>

# Our Motivation

<span style="color:#3399ff">
<font size="5">
Families from low-income communities struggle with literacy for several reasons. Limited education, a need to work multiple jobs, and little knowledge or access to resources. Arbolibro empowers and educates the community offering free literature for all ages.
</font>
</span>


<span style="color:#3399ff">
<font size="5">
In the United States, book-sharing has been promoted as a valuable activity that parents and children can do together to enhance literacy skills. Reading with your child has many positive benefits and is the key to your child’s academic future. 
</font>
</span>

```{r, echo=FALSE, fig.height=4, fig.width=4, fig.align='center'}
# image packages
library(png)
library(grid)
img <- readPNG('/Users/cristiannuno/Desktop/Syracuse/book3.png')
grid.raster(img)
```

<span style="color:#3399ff">
<font size="5">
It benefits children’s neurological, educational, psychological, social and linguistic future. Arbolibro offers a wide range of books from all ages in both English and Spanish. This project will bring an alternative to the community bringing them together.
</font>
</span>

# Our Locations

<span style="color:#3399ff">
<font size="5">
We currently have 4 locations in Little Village and Pilsen. We are looking forward to expanding to neighborhoods throughout Chicago in the near future!
</font>
</span>

```{r, echo=FALSE, fig.height=4, fig.width=8, fig.align='center'}
library( leaflet )
library( magrittr )
library( readxl )
arbolibro <- read_excel( path = "/Users/cristiannuno/Desktop/Syracuse/arbolibro.xlsx"
                   , sheet = 1
                   )
# Transform location urls from static to active
arbolibro$ActiveURL <- paste0("<a href='"
                                   , arbolibro$LocationWebsite
                                   , "' target='_blank'>"
                                   , arbolibro$LocationWebsite
                                   ,"</a>"
                                   )
# Build Map
mapArbolibro <- leaflet( data = arbolibro ) %>% # add locations dataset
  setView( lng = -87.705823, lat = 41.845816, zoom = 12) %>% # Set map to Little Village neighborhood
  addTiles() %>%  # Add map background
  addMarkers( lng = ~Long, lat = ~Lat
              # Add useful information onto each marker's popup
              , popup = paste0( "<b> Location Name: </b>"
                                , arbolibro$LocationName
                                , "<br>"
                                , "<b>Location Address: </b>"
                                , arbolibro$Location
                                , "<br>"
                                , "<b> Location Type: </b>"
                                , arbolibro$Type
                                , "<br>"
                                , "<b> Location Website: </b>"
                                , arbolibro$ActiveURL
                                )
              ) %>%
  addProviderTiles( "OpenStreetMap_DE" )

# Display Map
mapArbolibro

```

# In the News

<span style="color:#3399ff">
<font size="5">
Arbolibro has [received community attention](https://soundcloud.com/vocalo/domingosenvocalo-rayuelafm07316) from Lawndale News and Vocalo 91.1 FM, a segment called Rayuela FM which focuses on literature.
</font>
</span>


<span style="color:#3399ff">
<font size="3">
The interview starts at the 21:15 mark.
</font>
</span>


# Feedback

<span style="color:#3399ff">
<font size="5">
Please support the Arbolibro project by [recommending where where we should build our next Arbolibro](https://goo.gl/forms/yzEytmYNrLCtumaF2).
</font>
</span>


<span style="color:#3399ff">
<font size="5">
For immediate feedback, please email the founder of Arborlibro, Gonzalo Rizo, at gonzalorizo@hotmail.com. *This page was last updated on April 12, 2017*.
</font>
</span>

# Website Design

<span style="color:#3399ff">
<font size="5">
This website was designed by Cristian E. Nuno. To view more of his work, see his [GitHub page](https://github.com/cenuno) and his [LinkedIn page](https://www.linkedin.com/in/cristiannuno). You may email him at nuno.e.cristian@gmail.com.
</font>
</span>
